---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


We consider the problem of assessing the equality of multiple
high-dimensional (HD) populations or treatment outcomes, commonly
referred to as the Multivariate Analysis of Variance (MANOVA) problem.

Suppose the data involves $g$ independent groups, where $g\ge 2$, and
the $i$-th group consists of $n_i$ of $p$-variate independent and
identically distributed (i.i.d.) random vectors, denoted by
$$\pmb{X}_{ij}=(X_{ij1},\ldots,X_{ijp})^t,$$ for $i=1,\ldots,g, j=1,\ldots,n_i$.
Denote the $i$-th group mean vector by $\pmb{\mu}_{i}$. The MANOVA
testing problem considers the hypothesis:
$$H_0: \pmb{\mu}_1 = \cdots =\pmb{\mu}_g,$$
$$H_{a}: \pmb{\mu}_{i}\neq \pmb{\mu}_{j}, \text{ for at least on pair of }  i, j.$$

This package includes four testing procedures “SKtest” (Srivastava and
Kubokawa, 2013), “CXtest” (Cai and Xia, 2014)), “HBtest” (Hu et. al.,
2017), and “KVtest” (Kong, et. al., 2025). The input is a list of $g$
matrices, where the $i$-th matrix has dimension $(n_i\times p)$.

The function “rdatalist” is used to generate a list of sample data:
$$X_{ijk}=\mu_{ik}+\varepsilon_{ijk},\quad 1\le i \le g,\ 1\le j\le n_i,\ 1\le k\le p,$$
where $\mu_{ik}$ is the mean or location vectors, and the error
$\{\varepsilon_{ijk}: k = 1,\ldots, p\}$ is generated from various
innovation distributions

- **normal**: Normal(mean = 0, sd = 1);
- **t**: t(df = 3);
- **cauchy**: Cauchy(location = 0, scale = 0.1);
- **gamma**: centered Gamma(shape = 4, scale = 2).

and dependence structures:

- **Independent (IND)**: independent generation from the given
  innovation distribution.
- **Weakly Dependent (WD)**: random generation of ARMA(2, 2) model from
  the given innovation distribution, where autoregressive coefficients
  $\phi = (0.4, -0.1)$ and moving-average coefficients
  $\theta = (0.2, 0.3)$;
- **Strongly Dependent (SD)**: random generation of AR(1) model from the
  given innovation distribution, where autoregressive coefficients
  $\phi = 0.9$;
- **Long-range Dependent (LD)**:
  $\pmb{\varepsilon}_{ij}=U^t\pmb{\eta}_{ij}$, where $\pmb\eta_{ij}$ is
  $p$ independently generated from the given innovation distribution and
  $U^tU = A$ is the Cholesky factorization of the $p \times p$ matrix
  $A=(a_{st})$,
  $a_{st}=0.5[(|s-t|+1)^{1.4}+(|s-t|-1)^{1.4}-2|s-t|^{1.4}].$



```{r setup}
library(HDMANOVA)
set.seed(123)
X.list <- rdatalist(g = 3)
str(X.list)

SKtest(X.list)

CXtest(X.list)

CXtest(X.list, A = "Omega")

HBtest(X.list)

KVtest(X.list)

KVtest(X.list, method = "MPT", L = NA)

```
